---
title: "Candy Candy Candy"
output: html_notebook
---
Libraries required for this analysis
```{r}
library(tidyverse)
library(here)
```

load data
```{r}
candy <- read_csv(here("clean_data/clean_candy_data.csv"))
candy
```

#Q1 - What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

_Used !is.na to filter out NA values from the rating column_
_There are a total of 772352 ratings for candy across the three years_
_These are made up of 308182 "Joy" ratings, 90495 "Meh" ratings and 373675 "Despair" ratings_
```{r}

candy %>% 
  filter(!is.na(rating)) %>% 
  summarise(total_amount_of_all_ratings = n())

candy %>% 
  filter(!is.na(rating)) %>% 
  group_by(rating) %>% 
  summarise(total_of_ratings_by_grade = n())


```


#Q2 -What was the average age of people who are going out trick or treating?

_On first attempt the average age was 243 years old, need to clear up some outlier values in the age column - re-clean_
_After cleaning the mean age of tirck or treaters is 35_
_The median age of tirck or treaters is 37_
```{r}
candy %>% 
  filter(!is.na(age), trick_or_treating == "yes") %>%
  summarise(mean_age = mean(age))



candy %>% 
  filter(!is.na(age), trick_or_treating == "yes") %>%
  summarise(meadian_age = median(age))

```


#Q3 - What was the average age of people who are not going trick or treating?
_same method but different filter_
_The mean age not going trick or treating is 39_
_The median age not going trick or treating is 38_
```{r}
candy %>% 
  filter(!is.na(age), trick_or_treating == "no") %>%
  summarise(mean_age = mean(age))

candy %>% 
  filter(!is.na(age), trick_or_treating == "no") %>%
  summarise(median_age = median(age))
```


#Q4 - For each of joy, despair and meh, which candy bar received the most of these ratings?
_Below by filtering on the individual ratings and then grouping the most ratings are as follows;_
_joy = any_full_sized_candy_bar with 7589 ratings (obviously)_
_despair = anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes 10677 ratings_
_meh = lollipops 1570 ratings_
```{r}
candy %>% 
  filter(rating == "joy" & !is.na(candy_type)) %>% 
  group_by(candy_type, rating) %>% 
  summarise(number_of_joy_ratings = n()) %>% 
  ungroup() %>% 
  slice_max(number_of_joy_ratings)



candy %>% 
  filter(rating == "despair" & !is.na(candy_type)) %>% 
  group_by(candy_type, rating) %>% 
  summarise(number_of_despair_ratings = n()) %>% 
  ungroup() %>% 
  slice_max(number_of_despair_ratings)


candy %>% 
  filter(rating == "meh" & !is.na(candy_type)) %>% 
  group_by(candy_type, rating) %>% 
  summarise(number_of_meh_ratings = n()) %>% 
  ungroup() %>% 
  slice_max(number_of_meh_ratings)


#flipped rating and candy type around so that max for each rating was returned
candy %>%
  filter(!is.na(rating) & !is.na(candy_type)) %>% 
  group_by(rating, candy_type) %>% 
  summarise(number_of_ratings_for_candy_bar = n()) %>% 
  slice_max(number_of_ratings_for_candy_bar)
  


```

#Q5 - How many people rated Starburst as despair?
_Filtered by starburst and despair to find result_
_1990 ratings of despair were given to Starburst_

```{r}

candy %>% 
  filter(rating == "despair" & candy_type == "starburst") %>% 
  group_by(candy_type, rating) %>% 
  summarise(number_of_despair_ratings_for_starburst = n())

```

##For the next three questions, count despair as -1, joy as +1, and meh as 0.

_Used a case when function to add a new column and create a modified data set for these 3 questions_
```{r}


candy_rating_system <- candy %>% 
  mutate(
    candy_rating = case_when(
      rating == "despair" ~ -1,
      rating == "meh"     ~  0,
      rating == "joy"     ~  1
    )
  )

candy_rating_system
```


#Q6 - What was the most popular candy bar by this rating system for each gender in the dataset ?
_everybody loves full size candy bars (obviously)_

```{r}

candy_rating_system %>%
  filter(!is.na(rating)) %>% 
  group_by(gender, candy_type) %>% 
  summarise(popularity_score = sum(candy_rating)) %>% 
  slice_max(popularity_score)

#check using males data
candy_rating_system %>% 
  filter(gender == "Male" & !is.na(rating)) %>% 
  group_by(candy_type) %>% 
  summarise(popularity_score = sum(candy_rating)) %>% 
  ungroup() %>% 
  arrange(desc(popularity_score))
```



#Q7 - What was the most popular candy bar in each year?
_any_full_sized_candy_bar was the most popular candy bar every year (obviously)_
```{r}

candy_rating_system %>%
  filter(!is.na(rating)) %>% 
  group_by(year, candy_type) %>% 
  summarise(popularity_score = sum(candy_rating)) %>%
  slice_max(popularity_score)

```




#Q8 - What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

_Full sized candy bars are the most popular trick or treat item canada, usa and other countries. uk are more likely to be just after that straight cash money_
```{r}
candy_rating_system %>% 
  mutate(country_filtered = case_when(
    country == "usa" ~ "usa",
    country == "uk"  ~ "uk",
    country == "canada" ~ "canada",
    TRUE             ~ "other"
  )) %>% 
  filter(!is.na(rating)) %>% 
  group_by(country_filtered, candy_type) %>% 
  summarise(popularity_score = sum(candy_rating)) %>% 
  slice_max(popularity_score)
```

# Just for interest

__What was the least popular thing?__

The least popular by this candy rating system was the brown globs that had previously got the most despair ratings.


```{r}
candy_rating_system %>%
  filter(!is.na(rating)) %>% 
  group_by(candy_type) %>% 
  summarise(popularity_score = sum(candy_rating)) %>%
  slice_min(popularity_score)
```

__What is the most meh overall item to receive, i.e. the one whose candy rating is closest to zero while still getting a lot of ratings?__

The 3 most meh overall were mike and ike, mint m&ms and vicodin (which doesn't seem very suitable to give away as a trick or treat!).

```{r}
candy_rating_system %>%
  filter(!is.na(rating)) %>% 
  group_by(candy_type) %>% 
  summarise(popularity_score = sum(candy_rating), responses = n()) %>% 
  filter(popularity_score <100 & popularity_score >-100)
```

__What is the most and least popular candy by age group?__

Under 25 seem to enjoy cash as a trick or treat item while over 18s enjoy full size chocolate bars the most.

People up to 40 dislike the brown globs the most, after that age overall they are least satisfied with a broken glowstick (understandably).
```{r}
candy_rating_system %>%
  mutate(age_groups = case_when(
    age <18 ~ "under 18",
    age <26 ~ "18 to 25",
    age <41 ~ "26 to 40",
    age <61 ~ "41 to 60",
    age <105 ~ "over 60"
  )) %>% 
  filter(!is.na(rating)) %>% 
  group_by(age_groups, candy_type) %>% 
  summarise(popularity_score = sum(candy_rating)) %>% 
  slice_max(popularity_score)

candy_rating_system %>%
  mutate(age_groups = case_when(
    age <18 ~ "under 18",
    age <26 ~ "18 to 25",
    age <41 ~ "26 to 40",
    age <61 ~ "41 to 60",
    age <105 ~ "over 60"
  )) %>% 
  filter(!is.na(rating)) %>% 
  group_by(age_groups, candy_type) %>% 
  summarise(popularity_score = sum(candy_rating)) %>% 
  slice_min(popularity_score)
  
```

__What candy recived the most ratings? And how were the ratings distributed?__

The brown globs/mary janes re the most rated candy when ignoring NA values. This is partly because they were in some data sets multiple times, as both anonymous_brown_globs... and mary_janes.
```{r}

candy %>% 
  filter(!is.na(rating)) %>% 
  group_by(candy_type) %>% 
  summarise(total_ratings = n())  
```




